Best Practices for Multiple Alchemical Free Energy Calculations [Article v\versionnumber]

### 2025-05-26 paper/manuscript.tex
add citations

\paragraph{Perturbation maps}
\tocsectioncomment{Perturbation maps paragraph: @Mary Pitman/ @Jenke Scheen/ OpenFE team to update discussion of perturbation maps?}
For RBFE campaigns with more than two ligands in the input series a \textit{perturbation map} or \textit{-network} must be generated: this is simply a set of data that describe which transformations will be executed between the ligands (referred to as \textit{edges}). \cite{MOBLEY LOMAP} Heuristics have shown the more connected the perturbation network the higher the free energy prediction accuracy versus experimental measures. Running a fully-connected network (i.e. simulate transformations between \textit{all} ligands in the campaign) is typically considered too computationally expensive and may not be suited optimally with current estimators such as MLE (\ref{SECTION MLE}).\cite{SCHEEN ET AL} Several styles of network designs have been proposed in recent years, although no single method seems to consistently outperform all others.\cite{KARTOGRAF} There exist several methods for optimization of network structure while minimizing the number of perturbations that need to be computed, reducing the resulting computational cost at a minimal decrease in accuracy ~\cite{yang2020optimal,xu2019optimal}. Sometimes - depending on the RBFE implementation - the introduction of intermediate ligands that are not part of the original congeneric series can be helpful for avoiding challenging transformations such as ring breaking or transformations that would otherwise result in large numbers of atoms being inserted or deleted.\cite{LINDE PAPER} Some commercial tools have well-calibrated underlying heuristics to plan these network designs but may fail with complicated input datasets, thus requiring manual user validation, in particular when dealing with chiral compounds.

>>>>

\paragraph{Perturbation maps}
For RBFE campaigns with more than two ligands in the input series a \textit{perturbation map} or \textit{-network} must be generated: this is simply a set of data that describe which transformations will be executed between the ligands (referred to as \textit{edges}). ~\cite{liu2013lead} Heuristics have shown the more connected the perturbation network the higher the free energy prediction accuracy versus experimental measures. Running a fully-connected network (i.e. simulate transformations between \textit{all} ligands in the campaign) is typically considered too computationally expensive and may not be suited optimally with current estimators such as MLE (\ref{subsec:uncertainty}).~\cite{scheen22data} Several styles of network designs have been proposed in recent years, although no single method seems to consistently outperform all others.~\cite{ries2024kartograf} There exist several methods for optimization of network structure while minimizing the number of perturbations that need to be computed, reducing the resulting computational cost at a minimal decrease in accuracy ~\cite{yang2020optimal,xu2019optimal}. Sometimes - depending on the RBFE implementation - the introduction of intermediate ligands that are not part of the original congeneric series can be helpful for avoiding challenging transformations such as ring breaking or transformations that would otherwise result in large numbers of atoms being inserted or deleted.~\cite{schoenmaker2025imerge} Some commercial tools have well-calibrated underlying heuristics to plan these network designs but may fail with complicated input datasets, thus requiring manual user validation, in particular when dealing with chiral compounds.
